---
title: "httr Two Common Authentication Methods"
output: hugodown::md_document
status: "publish"
comment_status: "open"
categories: R Programming
tags:
  - httr
  - httr Authentication
  - httr API key
  - httr OAuth
---

What I just learned that there can be a couple of authentication methods used in R API wrapper packages but the most common ones are `API Key` and `OAuth 2.0`.

## API Key

`API Key` normally can be generated on a developer dashboard and can be regenerated later for security reason.

`API Key` can be saved as an environment variable in a `.Renviron` file and then we can use `Sys.getenv()` to get access to it, as shown here: <https://github.com/maelle/goodpress/blob/main/R/utils-http.R>.

`keyring` package is another common way to manage `API Key`: `key_set()` to create api key and `key_get()` to get access to the api key, as shown here: <https://github.com/lockedata/hubspot/blob/master/R/hubspot_key.R>.

Lastly, there are two different ways to include an api key in api calls: **including it as part of the URL or adding it in the HTTP header**.

`r goodpress::read_more()`

For example, HubSpot requires that api key needs to be added in the url (<https://developers.hubspot.com/docs/api/intro-to-auth>), while Leadfeeder asks developers to include it in the HTTP header (<https://docs.leadfeeder.com/api/#authentication>).

## OAuth 2.0

Take HubSpot for example. <https://developers.hubspot.com/docs/api/oauth-quickstart-guide> illustrates the steps in working with OAuth 2.0.

Accordingly, the steps in `httr` are:

-   **Step 1: provide the app information with `oauth_app()`**

```{r eval=FALSE}
oauth_app <- oauth_app(
  appname = app_name, 
  key = client_id, # from developer dashboard
  secret = client_secret # from developer dashboard
)
```

-   **Step 2: build an endpoint with `oauth_endpoint()`**

```{r eval=FALSE}
oauth_endpoint <- oauth_endpoint(
  authorize = authorize_url, # needs to be constructed first
  access = access_url # needs to be constructed first
)
```

-   **Step 3: create a token with `oauth2.0_token()`**

```{r eval=FALSE}
token <- oauth2.0_token(
  endpoint = oauth_endpoint,
  app = oauth_app
)
```

-   **Step 4: save the token into a file and create an environment variable pointing to it**

-   **Step 5: use the token in every api call**

Token can be accessed with `token$credentials$access_token` and should be added in HTTP headers with `add_headers()` or `config()`.

-   **Step 6: refresh token with `token$refresh()`**

Here is an example that implements all the steps above for HubSpot OAuth 2.0: <https://github.com/lockedata/hubspot/blob/master/R/hubspot_oauth.R>.

Besides learning from the source code on github, I found the following articles very useful:

-   [An Introduction to APIs](https://zapier.com/learn/apis/)
-   [Best practices for API packages](https://httr.r-lib.org/articles/api-packages.html)
