---
title: "Transform List into Dataframe With tidyr and purrr"
output: hugodown::md_document
status: "publish"
comment_status: "open"
categories: R Programming
tags:
  - List Manipulation in R
  - hoist() vs map()
  - list with tidyr vs purrr
---

As a data structure in R, list is not as familiar to me as vector and dataframe. I knew that list is often returned by function calls and I didn't pay much attention to it until I started working on the API wrapper package [RLeadfeeder](https://github.com/henrywangnl/RLeadfeeder). It turned out that list can be very useful to hold all kinds of data returned from API platforms and I had to make an effort to learn how to work with it, namely extract useful elements from a list and turn them into a dataframe.

This post is an example of how to transfrom a list into a dataframe with two different approaches: tidyr and purrr. The packages used in this post are as follows:

`r goodpress::read_more()`

```{r message=FALSE}
library(tidyr) # for hoist() and unnest_longer()
library(magrittr) # for extract()
library(purrr) # for map()
library(listviewer) # for jsonedit()
library(repurrrsive) # for gh_repos dataset
```


## Inspection

Start with the inspection of list elements using `listviewer` package.

**Task 1: Examine and understand list elements interactively.**

```{r eval=FALSE}
jsonedit(gh_repos)
```

## Extract Multiple Elements at the same level

**Task 2: Extract each repository's name and full name.**

### purrr approach

```{r eval=FALSE}
gh_repos %>% 
  map_df(~map(.x, `[`, c("name", "full_name")))
```


```{r}
# replace `[` with magrittr::extract() 
gh_repos %>% 
  map_df(~map(.x, magrittr::extract, c("name", "full_name")))
```

### tidyr approach

```{r}
tibble(repo = gh_repos) %>% 
  unnest_longer(repo) %>% 
  hoist(repo, "name", "full_name")
```


## Extract Multiple Elements at different levels

**Task 3: Extract each repository's name, full name, and owner's username (`owner` -> `login`).**

### purrr approach

```{r}
name <- gh_repos %>% 
  map(~map_chr(.x, "name")) %>% 
  flatten_chr()

full_name <- gh_repos %>% 
  map(~map_chr(.x, "full_name")) %>% 
  flatten_chr()

username <- gh_repos %>% 
  map(~map_chr(.x, c("owner", "login"))) %>% 
  flatten_chr()

tibble(name, full_name, username)
```

### tidyr approach 

```{r}
tibble(repo = gh_repos) %>% 
  unnest_longer(repo) %>% 
  hoist(repo, "name", "full_name",
        username = c("owner", "login"))
```

## Preference

As shown above, both approaches work fine but `tidyr` approach seems easier and more flexible to me. For example, considering the following question:

**Task 4: Extract the full name of each user's first repository.**

```{r}
tibble(repo = gh_repos) %>% 
  hoist(repo, 
        full_name = c(1, 3))
```

```{r}
tibble(repo = gh_repos) %>% 
  hoist(repo, 
        full_name = list(1, "full_name"))
```

## Resources

Last but not least I find these two resources are very useful for me to learn how to work with list:

* purrr tutorial: https://jennybc.github.io/purrr-tutorial/index.html
* tidyr rectangling: https://tidyr.tidyverse.org/articles/rectangle.html